---
title: "New Haven GSI Stewardship"
author: "Clare Auld-Brokish (with Dexter Locke)"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---


# 0 set up
```{r}

# Load libraries 
packs <-c('tidyverse'   # cuz
          , 'tidylog'   # prints out what was done in dplyr and tidyr; VERBOSE
          # , 'tidygeocoder' # geocode addressses
          , 'tidycensus'   # access Census data
          , 'sf'           # for spatial data support
          , 'mapview'      # web maps for zooming and panning around
          , 'janitor'   # helps clean things pipe-friendly cross-tabulations
          , 'leafsync'  # linked maps
          )

if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load the packages all at once
vapply(packs, library, character.only = TRUE, logical(1),
       logical.return = TRUE, quietly = TRUE)


```


# 1 Read in data
## A Neighborhoods
```{r}

neighs <- 
   st_read('input_data/New_Haven_Neighborhoods/Neighborhood.shp', as_tibble = TRUE) |> 
   select(Neighborhood = Neighbor_1) |> 
   st_transform(3857) %>%
   mutate(area_m2=as.double(st_area(.)))

neighs |> mapview(zcol = "area_m2")

```

## B Watersheds and Zoning
```{r}

sheds <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/12 digit Hydrologic Unit.geojson') |> 
  st_transform(st_crs(neighs))

mapview(neighs, alpha.regions = 0, lwd = 2) + mapview(sheds, zcol = 'Name') 

# st_layers('../GSI STEWARDSHIP Project_data_too_big/_ags_data0AA96A07DF994B819C3A75306E2DD073.gdb')

zoning <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/_ags_data0AA96A07DF994B819C3A75306E2DD073.gdb') |> 
  st_transform(st_crs(neighs)) |> 
  select(-shape_Length,-shape_Area) %>%
  mutate(area_m2=as.double(st_area(.)))

zoning |> glimpse()
zoning |> mapview(zcol = 
                    # 'zone_code'
                    'districts'
                    # 'district_category'
                  )

#How much area is in each zone?
zoning |> 
  st_drop_geometry() |> 
  group_by(districts) |> 
  summarize(total_area=sum(area_m2)
            ,n=n()
            , mean_area=mean(area_m2)
            , median_area=median(area_m2)
            , sd_area=sd(area_m2)) |> 
  arrange(desc(total_area)) 

```

# 2 Data Transformation
## A Inlet
```{r}
# Join neighborhoods, watersheds, and zoning in one chart called Inlet
inlet <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'inlet'
          , as_tibble = TRUE) |> 
  st_intersection(neighs) |> # adds neighborhood name
  st_intersection(sheds |> 
                    select(watershed = Name, HUC12)
                    ) |> 
  st_intersection(zoning |> 
                    select(zone_code, districts, district_category)
                  ) |> 
  mutate(  edit_year = lubridate::year(EditDate)
         , edit_month = lubridate::month(EditDate)
         , edit_date = lubridate::date(EditDate)
         , sun_vs_shade = str_extract(Planting_Plan, "^[^ ]+") |> str_to_title()
  ) 

# double check sun_vs_shade variable
inlet |>
  st_drop_geometry() |> 
  tabyl(Planting_Plan, sun_vs_shade)

# MapView
mapviewOptions(
    legend.pos = 'bottomleft'
  , layers.control.pos = 'topright'
  )

# View all the inlets at once by neighborhood, zone, and watershed 
m_neigh <- inlet |> mapview(zcol = 'Neighborhood') # TK
m_water <- inlet |> mapview(zcol = 'watershed')
m_zone  <- inlet |> mapview(zcol = 'districts')

leafsync::sync(m_neigh, m_water, m_zone)

```

### 1 Analysis
```{r}
# What is the distribution of planting plans across the city? 
## Percent of inlets with sun or shade planting plans
inlet |> 
  st_drop_geometry() |>
  tabyl(sun_vs_shade) |> 
  ggplot(aes(sun_vs_shade, percent)) +
  geom_col() + 
  theme_linedraw()

## Planting plans by watershed, neighborhood, or business district?

## What is the distribution of planting plans assigned "emergency cleaning" vs "routine cleaning"?

```

## B Inspection
```{r}
# Create dataset with inspection results, beginning with condition 
inspection <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'GI_Inspection'
          , as_tibble = TRUE) |> 
mutate(Condition=factor(CONDITION,
    levels=c("Unknown", NA, "Very Poor", "Poor", "Fair", "Good", "Very Good", "Excellent"))
  , islegit=ifelse(Condition=="Unknown" | is.na(Condition)
                   , "notlegit"
                   , "legit"), 
  edit_year = lubridate::year(EditDate), 
    edit_month = lubridate::month(EditDate),
    edit_date = lubridate::date(EditDate)
  ) |>
  select(-EditDate) 
  
# Double-checks classifications
inspection |> tabyl(Condition,islegit)

# This counts "missing-ness"
inspection |> map(~sum(is.na(.))) |> bind_rows() |> t()

# is the unique ID really unique? 
all.equal(
  inspection |> distinct(GlobalID) |> nrow()
, inspection |> nrow()
)

```

### B1 GSI per Neighborhood
```{r}
# Account for mean area of a Neighborhood when comparing inlets across neighborhoods 
gsi_per_neigh <- 
  inlet |> 
  ungroup() |> 
  st_drop_geometry() |> 
  group_by(Neighborhood) |> 
  count()

neighs |> 
  left_join(gsi_per_neigh, by = 'Neighborhood') |> 
  mutate(gsi_per_neigh_area = ifelse(is.na(n), 0, n / area_m2)) |> # we are accounting for NA's
  mapview(zcol = 'gsi_per_neigh_area', layer.name = 'gsi_per_neigh_area')

neighs |> 
  left_join(gsi_per_neigh, by = "Neighborhood") %>%
  mutate(n = ifelse(is.na(n), 0, n)
         , area_m2 = st_area(.)
         , gsi_per_area = n / area_m2
         ) |> 
  mapview(zcol = 'gsi_per_area', layer.name = 'GSI counts /<br>neigh area')

```

### B2 Inspection Frequencies
```{r}

inspection |> map(~sum(is.na(.))) |> bind_rows() |> t()

# Tabulate frequency of the asset_GLOBALID A
inspection_freq <- 
  inspection |> 
  tabyl(ASSETGUID) |> 
  as_tibble() |> 
  arrange(desc(n))

# How often have inlets been inspected?
inspection_freq |> 
  summary() # roughly 4x

# Valid percent of frequency distribution
inspection_freq |> 
  ggplot(aes(n, valid_percent)) + 
  geom_col() +
  theme_bw(16) +
  NULL 

# What's the distribution of frequencies?
inspection_freq_freq <- 
  inspection_freq |> 
  tabyl(n) |> 
  as_tibble()

inspection_freq_freq |> 
  ggplot(aes(n,n_n)) + 
  geom_col() +
  theme_bw(16) +
  NULL 

```

## C Recent_Inspections
```{r}
# Includes only the most recently performed inspection for each inlet
  
recent_inspections <- 
  inspection |> 
  group_by(ASSETGUID) |> 
  arrange(edit_date) |> 
  slice(1) |> # slice_max, slice_min instead of arrange |> slice 
  ungroup()
  
recent_inspections |> 
  tabyl(CONDITION) |> 
  as_tibble()

# Most recent inspections are poor, fair, or good 
recent_inspections |> 
  tabyl(Condition) |> 
  ggplot(aes(Condition, n)) +
  geom_col()

```

### 1 Analysis 
```{r}

# Most recent inspection for each inlet by year
recent_inspections |> 
  tabyl(edit_date)  |> 
  as_tibble() |> 
  ggplot(aes(edit_date, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth()

# Most recent inspection for each inlet by month
# The most recent inspection for each inlet has occurred in April and May 
recent_inspections |> 
  tabyl(edit_month)  |> 
  as_tibble() |> 
  ggplot(aes(edit_month, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth()

# Islegit vs Notlegit in Recent_inspections
recent_inspections |> 
  tabyl(Condition, islegit) |> 
  as_tibble() |> 
  pivot_longer(-Condition, names_to = "islegit") |> 
  ggplot(aes(value, Condition)) +
  geom_col() + 
  facet_wrap(~islegit, scales="free") + 
  NULL
```

### 2 Analysis - Debris, Vegetation, Oil
```{r}
# Debris - upon inspection, most inlets had debris
recent_inspections |> 
  tabyl(DEBRIS) |> 
  as_tibble() |> 
  arrange(desc(n))
  
recent_inspections |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = DEBRIS)) + 
  geom_col() + 
  theme_linedraw()

# The only bioswales with debris in them were assessed in a "poor", "very poor", and "NA" condition 
recent_inspections |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = DEBRIS)) +
  geom_col() +
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~DEBRIS) +
  NULL

# Vegetation - no bioswales with vegetation? 
recent_inspections |> 
  group_by(VEGETATION, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = VEGETATION)) +
  geom_col() + 
  NULL

# Oil - most bioswales did not have oil in them 
recent_inspections |> 
  group_by(Condition, OIL) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = OIL)) +
  geom_col() + 
  NULL

# Frame condition
recent_inspections |> 
  group_by(FRAME, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition)) +
  geom_col() + 
  facet_wrap(~FRAME, scales="free") + 
  NULL

```

## D Inspection_Long
```{r}
# Join inspections and inlets by FACILITYID and GlobalID to create a dataset that includes condition, neighborhood, watershed, and zone

inspection_long <- 
  inspection |> 
  left_join(
    inlet |> select(-c(CreationDate, edit_date, Creator, Editor, edit_year, edit_month, GlobalID))
              , by = c('FACILITYID' = 'FACILITYID'))

# Calculate the number of GSI inlets per district
n_gsi_district <- 
  inlet |> 
  group_by(districts) |> 
  summarize(  #total_area  = sum(area_m2)
            , n           = n()
            ) |> 
  st_drop_geometry()

zoning |> 
  group_by(districts) |> 
  summarise(district_area_m2 = sum(area_m2)) |> 
  left_join(n_gsi_district, by = 'districts') |> 
  mutate(gsi_per_district_area = ifelse(is.na(n), 0, n/district_area_m2))

```

### 1 Analysis - Total Inspections
```{r}
# Inspections by watershed per district 
inspection_long |> 
  group_by(districts, watershed) |> 
  count(districts) |> 
  ggplot(aes(districts, n)) +
  geom_col() + 
  facet_wrap(~watershed) + 
  theme(axis.text.x=element_text(angle = 90)) 

# Inspections focus primarily on the Cove River-Frontal LI Sound watershed
inspection_long |> 
  group_by(watershed) |> 
  count() |> 
  ggplot(aes(watershed, n)) + # 'n' was missing
  geom_col() + 
  theme_classic()

# Inspections focus primarily on the Cove River-Frontal LI Sound watershed in the Hill Neighborhood
inspection_long |> 
  group_by(watershed, Neighborhood) |> 
  count() |> 
  ggplot(aes(n, watershed)) +
  geom_col() + 
  theme_linedraw() + 
  facet_wrap(~Neighborhood)

```

### 2 Analysis - Cleaning Assignments 
```{r}
# What is in the categories of assigned?
inspection_long |> 
  tabyl(Assigned) |> 
  as_tibble()

# Assigned by watershed, neighborhood, or district 
inspection_long |> 
  tabyl(Neighborhood, Assigned) |> 
  as_tibble()

# Most inlets are assigned emergency cleaning
inspection_long |> 
  tabyl(Assigned) |> 
  ggplot(aes(n, Assigned)) +
  geom_col() + 
  theme_linedraw() + 
  NULL 

# Inlet condition by cleaning assignments 
# The assignment of cleaning does not measurably impact inlet condition
inspection_long |> 
  group_by(Assigned, Condition) |> 
  count('Routine Cleaning') |> 
  ggplot(aes(n, Condition)) + 
  geom_col() + 
  facet_wrap(~Assigned, scales="free") +
NULL

# Inlet condition by cleaning assignment and neighborhood
# Inlets in the Hill Neighborhood have the highest number of inspections and emergency cleanings
inspection_long |> 
  group_by(Neighborhood, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~Neighborhood)

# Inlet condition by cleaning assignment and watershed
# Characterization of the cleanings in the Cove River-Frontal LI Sound watershed
inspection_long |> 
  group_by(watershed, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~watershed)

```

### 3 Analysis - Condition 
```{r}
# Inlet condition by watershed and neighborhood
inspection_long |> 
  ggplot(aes(Condition, fill = Neighborhood)) +
  geom_bar() +
  facet_wrap(~watershed, scales = "free_y") + 
  theme(axis.text.x=element_text(angle = 90)) +
NULL 

# Inlet condition by edit month 
inspection_long |> 
  group_by(edit_month, Condition) |> 
  summarize(n=n()) |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_bw(16) + 
  facet_wrap(~edit_month, nrow=3) + 
  NULL

# Inlet condition by presence of debris 
# There is no discernible correlation between inlet condition and the presence of debris
inspection_long |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  facet_wrap(~DEBRIS) + 
  NULL

```

## E Inlet_Recent_Inspections
```{r}
# Combines neighborhood, watershed, and district data with recent inspections data

inlet_recent_inspections <- 
  inlet |> 
  left_join(
    recent_inspections |> select(GlobalID, ASSETGUID, FACILITYID, DEBRIS, VEGETATION, OIL, FRAME, Condition, islegit)
    , by = c('GlobalID' = 'ASSETGUID'))

```

### 1 Analysis 
```{r}
inlet_recent_inspections |> 
  st_drop_geometry() |> 
  group_by(Condition, islegit) |> 
  count()

# Inlet condition by watershed
inlet_recent_inspections |> 
  group_by(watershed, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~watershed) +
  NULL

# Inlet condition by district 
inlet_recent_inspections |> 
  group_by(districts, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~districts, scales="free") +
  NULL

# Inlet condition by district and watershed
inlet_recent_inspections |> 
  group_by(watershed, districts, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = districts)) +
  geom_col() + 
  facet_wrap(~watershed, scales="free") +
  theme(axis.text.x=element_text(angle = 90)) + 
  NULL

# Most inspections occur April through July (by district category)
inlet_recent_inspections |> 
  group_by(edit_month, Condition, district_category) |> 
  summarize(n=n()) |> 
  ggplot(aes(district_category, n, fill = Condition)) +
  geom_col() +
  theme_bw(10) + 
  facet_wrap(~edit_month, nrow=3) + 
  theme(axis.text.x=element_text(angle = 90)) +
  NULL

# The least number of inspections occur in residential districts
inlet_recent_inspections |> 
  group_by(district_category) |> 
  count()|> 
  ggplot(aes(district_category, n)) +
  geom_col() +
  theme_bw(10) +
  theme(axis.text.x=element_text(angle = 90)) + 
  NULL

# What is the condition of recent inspections by district category? 
inlet_recent_inspections |> 
  group_by(district_category, Condition) |> 
  count()|> 
  ggplot(aes(district_category, n, fill = Condition)) +
  geom_col() +
  theme_bw(10) +
  theme(axis.text.x=element_text(angle = 90))

# Most frame repair inspections occur in high-middle density districts
inlet_recent_inspections |> 
  group_by(FRAME, district_category) |> 
  count() |> 
  ggplot(aes(n, district_category)) +
  geom_col() + 
  theme_bw(10) + 
  facet_wrap(~FRAME) + 
  NULL

# Inlet condition by neighborhood
# Inlet condition by neighborhood, districts, and district categories
# When do inspections occur (month, year) by neighborhood, districts, and district categories?
# When do inspections occur (month, year) by neighborhood, districts, cleaning type, and sun_vs_shade? 

```

## F Parks
```{r}
(parks <- st_read('input_data/NewHavenParksLayers_08032021/NewHavenParks.shp'))

parks |> mapview()

mapview(parks) +
  mapview(neighs, zcol = "Neighborhood", layer.name = "neighborhood") +
  mapview(inlet)

```

## G Parcels
```{r}
```



