---
title: "New Haven GSI Stewardship"
author: "Clare Auld-Brokish (with Dexter Locke)"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---


# 0 set up
```{r echo=FALSE}

# Load libraries 
packs <-c('tidyverse'   # cuz
          , 'tidylog'   # prints out what was done in dplyr and tidyr; VERBOSE
          # , 'tidygeocoder' # geocode addressses
          , 'tidycensus'   # access Census data
          , 'sf'           # for spatial data support
          , 'mapview'      # web maps for zooming and panning around
          , 'janitor'   # helps clean things pipe-friendly cross-tabulations
          , 'leafsync'  # linked maps
          )

if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load the packages all at once
vapply(packs, library, character.only = TRUE, logical(1),
       logical.return = TRUE, quietly = TRUE)


```


# 1 Read in data
## A Neighborhoods
```{r echo=FALSE}

neighs <- 
   st_read('input_data/New_Haven_Neighborhoods/Neighborhood.shp', as_tibble = TRUE) |> 
   select(Neighborhood = Neighbor_1) |> 
   st_transform(3857) %>%
   mutate(area_m2=as.double(st_area(.)))

neighs |> mapview(zcol = "area_m2")

```

## B Watersheds and Zoning
```{r echo=FALSE}

sheds <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/12 digit Hydrologic Unit.geojson') |> 
  st_transform(st_crs(neighs))

mapview(neighs, alpha.regions = 0, lwd = 2) + mapview(sheds, zcol = 'Name') 

# st_layers('../GSI STEWARDSHIP Project_data_too_big/_ags_data0AA96A07DF994B819C3A75306E2DD073.gdb')

zoning <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/_ags_data0AA96A07DF994B819C3A75306E2DD073.gdb') |> 
  st_transform(st_crs(neighs)) |> 
  select(-shape_Length,-shape_Area) %>%
  mutate(area_m2=as.double(st_area(.)))

zoning |> glimpse()
zoning |> mapview(zcol = 
                    # 'zone_code'
                    'districts'
                    # 'district_category'
                  )

#How much area is in each zone?
zoning |> 
  st_drop_geometry() |> 
  group_by(districts) |> 
  summarize(total_area=sum(area_m2)
            ,n=n()
            , mean_area=mean(area_m2)
            , median_area=median(area_m2)
            , sd_area=sd(area_m2)) |> 
  arrange(desc(total_area)) 

```

# 2 Data Transformation
## A Inlet
Join neighborhoods, watersheds, and zoning in one dataset called Inlet
```{r echo=FALSE}
inlet <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'inlet'
          , as_tibble = TRUE) |> 
  st_intersection(neighs) |> # adds neighborhood name
  st_intersection(sheds |> 
                    select(watershed = Name, HUC12)
                    ) |> 
  st_intersection(zoning |> 
                    select(zone_code, districts, district_category)
                  ) |> 
  mutate(  edit_year = lubridate::year(EditDate)
         , edit_month = lubridate::month(EditDate)
         , edit_date = lubridate::date(EditDate)
         , sun_vs_shade = str_extract(Planting_Plan, "^[^ ]+") |> str_to_title()
  ) 

```

Double check sun_vs_shade variable
```{r echo=FALSE}
inlet |>
  st_drop_geometry() |> 
  tabyl(Planting_Plan, sun_vs_shade)

# MapView
mapviewOptions(
    legend.pos = 'bottomleft'
  , layers.control.pos = 'topright'
  )

```

View all inlets at once by neighborhood, zone, and watershed
``` {r echo=FALSE}
m_neigh <- inlet |> mapview(zcol = 'Neighborhood') # TK
m_water <- inlet |> mapview(zcol = 'watershed')
m_zone  <- inlet |> mapview(zcol = 'districts')

leafsync::sync(m_neigh, m_water, m_zone)

```

### 1 Analysis
What is the distribution of planting plans (sun vs. shade) across the city?
```{r echo=FALSE}
inlet |> 
  st_drop_geometry() |>
  tabyl(sun_vs_shade) |> 
  ggplot(aes(sun_vs_shade, percent)) +
  geom_col() + 
  theme_linedraw()
```

Remaining questions/potential graphs: 
- Planting plans by watershed, neighborhood, or business district?
- What is the distribution of planting plans assigned "emergency cleaning" vs "routine cleaning"?

## B Inspection
Create a dataset with inspection results, beginning with condition
```{r echo=FALSE}
inspection <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'GI_Inspection'
          , as_tibble = TRUE) |> 
mutate(Condition=factor(CONDITION,
    levels=c("Unknown", NA, "Very Poor", "Poor", "Fair", "Good", "Very Good", "Excellent"))
  , islegit=ifelse(Condition=="Unknown" | is.na(Condition)
                   , "notlegit"
                   , "legit"), 
  edit_year = lubridate::year(EditDate), 
    edit_month = lubridate::month(EditDate),
    edit_date = lubridate::date(EditDate)
  ) |>
  select(-EditDate) 

```

Double-check classifications
``` {r echo=FALSE}
# Double-checks classifications
inspection |> tabyl(Condition,islegit)

```

Count "missing-ness" 
```{r echo=FALSE}
inspection |> map(~sum(is.na(.))) |> bind_rows() |> t()

```

Is the unique ID really unique? 
```{r echo=FALSE}
all.equal(
  inspection |> distinct(GlobalID) |> nrow()
, inspection |> nrow()
)

```

### B1 GSI per Neighborhood
Account for mean area of a neighborhood when comparing inlets across neighborhoods
```{r echo=FALSE}
gsi_per_neigh <- 
  inlet |> 
  ungroup() |> 
  st_drop_geometry() |> 
  group_by(Neighborhood) |> 
  count()

neighs |> 
  left_join(gsi_per_neigh, by = 'Neighborhood') |> 
  mutate(gsi_per_neigh_area = ifelse(is.na(n), 0, n / area_m2)) |> # we are accounting for NA's
  mapview(zcol = 'gsi_per_neigh_area', layer.name = 'gsi_per_neigh_area')

neighs |> 
  left_join(gsi_per_neigh, by = "Neighborhood") %>%
  mutate(n = ifelse(is.na(n), 0, n)
         , area_m2 = st_area(.)
         , gsi_per_area = n / area_m2
         ) |> 
  mapview(zcol = 'gsi_per_area', layer.name = 'GSI counts /<br>neigh area')

```

### B2 Inspection Frequencies
```{r echo=FALSE}

inspection |> map(~sum(is.na(.))) |> bind_rows() |> t()

```

Tabulate frequency of the asset_GLOBALID
```{r echo=FALSE}
inspection_freq <- 
  inspection |> 
  tabyl(ASSETGUID) |> 
  as_tibble() |> 
  arrange(desc(n))

```

How often have inlets been inspected? 
```{r echo=FALSE}
inspection_freq |> 
  summary() # roughly 4x

```
Analysis: Roughly four times. 

Valid percent of frequency distribution
```{r echo=FALSE}
inspection_freq |> 
  ggplot(aes(n, valid_percent)) + 
  geom_col() +
  theme_bw(16) +
  NULL 

```

What is the distribution of frequencies? 
```{r echo=FALSE}
inspection_freq_freq <- 
  inspection_freq |> 
  tabyl(n) |> 
  as_tibble()

inspection_freq_freq |> 
  ggplot(aes(n,n_n)) + 
  geom_col() +
  theme_bw(16) +
  NULL 

```

## C Recent_Inspections
Includes only the most recently performed inspection for each inlet
```{r echo=FALSE}
recent_inspections <- 
  inspection |> 
  group_by(ASSETGUID) |> 
  arrange(edit_date) |> 
  slice(1) |> # slice_max, slice_min instead of arrange |> slice 
  ungroup()
  
recent_inspections |> 
  tabyl(CONDITION) |> 
  as_tibble()

```

Recent inspections by condition
```{r echo=FALSE}
recent_inspections |> 
  tabyl(Condition) |> 
  ggplot(aes(Condition, n)) +
  geom_col()

```
Analysis: Most recent inspections are poor, fair, or good

### 1 Analysis 
Most recent inspection for each inlet by year 
```{r echo=FALSE}
recent_inspections |> 
  tabyl(edit_date)  |> 
  as_tibble() |> 
  ggplot(aes(edit_date, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth()

```

Most recent inspection for each inlet by month
```{r echo=FALSE}
recent_inspections |> 
  tabyl(edit_month)  |> 
  as_tibble() |> 
  ggplot(aes(edit_month, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth()

```
Analysis: The most recent inspection for each inlet occurred in April and May.

Islegit vs. Notlegit in recent inspections
```{r echo=FALSE}
recent_inspections |> 
  tabyl(Condition, islegit) |> 
  as_tibble() |> 
  pivot_longer(-Condition, names_to = "islegit") |> 
  ggplot(aes(value, Condition)) +
  geom_col() + 
  facet_wrap(~islegit, scales="free") + 
  NULL
```

### 2 Analysis - Debris, Vegetation, Oil
Debris
```{r echo=FALSE}
recent_inspections |> 
  tabyl(DEBRIS) |> 
  as_tibble() |> 
  arrange(desc(n))
  
recent_inspections |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = DEBRIS)) + 
  geom_col() + 
  theme_linedraw()

```
Analysis: Most bioswales had debris. 

```{r echo=FALSE}
recent_inspections |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = DEBRIS)) +
  geom_col() +
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~DEBRIS) +
  NULL
```
Analysis: The only bioswales with debris in them were assessed in a "poor", "very poor", and "NA" condition.

Vegetation
```{r echo=FALSE}
recent_inspections |> 
  group_by(VEGETATION, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = VEGETATION)) +
  geom_col() + 
  NULL

```
Analysis: Most bioswales did not have vegetation.

Oil 
```{r echo=FALSE}
recent_inspections |> 
  group_by(Condition, OIL) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = OIL)) +
  geom_col() + 
  NULL

```
Analysis: Most bioswales did not have oil. 

Frame condition
```{r echo=FALSE}
recent_inspections |> 
  group_by(FRAME, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition)) +
  geom_col() + 
  facet_wrap(~FRAME, scales="free") + 
  NULL

```

## D Inspection_Long
Join inspections and inlets by FACILITYID and GlobalID to create a dataset that includes condition, neighborhood, watershed, and zone
```{r echo=FALSE}
inspection_long <- 
  inspection |> 
  left_join(
    inlet |> select(-c(CreationDate, edit_date, Creator, Editor, edit_year, edit_month, GlobalID))
              , by = c('FACILITYID' = 'FACILITYID'))

```

Calculate the number of inlets per district
```{r echo=FALSE}
n_gsi_district <- 
  inlet |> 
  group_by(districts) |> 
  summarize(  #total_area  = sum(area_m2)
            , n           = n()
            ) |> 
  st_drop_geometry()

zoning |> 
  group_by(districts) |> 
  summarise(district_area_m2 = sum(area_m2)) |> 
  left_join(n_gsi_district, by = 'districts') |> 
  mutate(gsi_per_district_area = ifelse(is.na(n), 0, n/district_area_m2))

```

### 1 Analysis - Total Inspections
Inspections by watershed 
```{r echo=FALSE}
inspection_long |> 
  group_by(watershed) |> 
  count() |> 
  ggplot(aes(watershed, n)) + # 'n' was missing
  geom_col() + 
  theme_classic()

```
Analysis: Inspections focus primarily on the Cove River-Frontal LI Sound watershed

Inspections by watershed and district
```{r echo=FALSE}
inspection_long |> 
  group_by(districts, watershed) |> 
  count(districts) |> 
  ggplot(aes(districts, n)) +
  geom_col() + 
  facet_wrap(~watershed) + 
  theme(axis.text.x=element_text(angle = 90)) 

```

Inspections by watershed and neighborhood
```{r echo=FALSE}
inspection_long |> 
  group_by(watershed, Neighborhood) |> 
  count() |> 
  ggplot(aes(n, watershed)) +
  geom_col() + 
  theme_linedraw() + 
  facet_wrap(~Neighborhood)

```
Analysis: Inspections focus primarily on the Cove River-Frontal LI Sound watershed in the Hill Neighborhood

### 2 Analysis - Cleaning Assignments 
What is in the categories of assigned?
```{r echo=FALSE}
inspection_long |> 
  tabyl(Assigned) |> 
  as_tibble()

```

Assigned by watershed, neighborhood, or district
```{r echo=FALSE}
inspection_long |> 
  tabyl(Neighborhood, Assigned) |> 
  as_tibble()

```

```{r echo=FALSE}
inspection_long |> 
  tabyl(Assigned) |> 
  ggplot(aes(n, Assigned)) +
  geom_col() + 
  theme_linedraw() + 
  NULL 

```
Analysis: Most inlets are assigned emergency cleaning. 

Inlet condition by cleaning assignments 
```{r echo=FALSE}
inspection_long |> 
  group_by(Assigned, Condition) |> 
  count('Routine Cleaning') |> 
  ggplot(aes(n, Condition)) + 
  geom_col() + 
  facet_wrap(~Assigned, scales="free") +
NULL

```
Analysis: The assignment of cleaning does not measurably impact inlet condition.

Inlet condition by cleaning assignment and neighborhood
```{r echo=FALSE}
inspection_long |> 
  group_by(Neighborhood, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~Neighborhood)

```
Analysis: Inlets in the Hill Neighborhood have the highest number of inspections and emergency cleanings

Inlet condition by cleaning assignment and watershed 
```{r echo=FALSE}
inspection_long |> 
  group_by(watershed, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~watershed)

```

### 3 Analysis - Condition 
Inlet condition by watershed and neighborhood
```{r echo=FALSE}
inspection_long |> 
  ggplot(aes(Condition, fill = Neighborhood)) +
  geom_bar() +
  facet_wrap(~watershed, scales = "free_y") + 
  theme(axis.text.x=element_text(angle = 90)) +
NULL 

```

Inlet condition by edit month
```{r echo=FALSE}
inspection_long |> 
  group_by(edit_month, Condition) |> 
  summarize(n=n()) |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_bw(16) + 
  facet_wrap(~edit_month, nrow=3) + 
  NULL

```

Inlet condition by presence of debris
```{r echo=FALSE}
inspection_long |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  facet_wrap(~DEBRIS) + 
  NULL

```
Analysis: There is no discernible correlation between inlet condition and the presence of debris.

## E Inlet_Recent_Inspections
Combines neighborhood, watershed, and district data with recent inspections data
```{r echo=FALSE}
inlet_recent_inspections <- 
  inlet |> 
  left_join(
    recent_inspections |> select(GlobalID, ASSETGUID, FACILITYID, DEBRIS, VEGETATION, OIL, FRAME, Condition, islegit)
    , by = c('GlobalID' = 'ASSETGUID'))

```

### 1 Analysis 
```{r echo=FALSE}
inlet_recent_inspections |> 
  st_drop_geometry() |> 
  group_by(Condition, islegit) |> 
  count()

```

Inlet condition by watershed
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(watershed, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~watershed) +
  NULL

```

Inlet condition by district
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(districts, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  theme_bw(16) +
  theme(legend.position = "bottom") +
  facet_wrap(~districts, scales="free") +
  NULL

```

Inlet condition by district and watershed
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(watershed, districts, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = districts)) +
  geom_col() + 
  facet_wrap(~watershed, scales="free") +
  theme(axis.text.x=element_text(angle = 90)) + 
  NULL

```

Inlet condition by edit month and district category
```{r echo=FALSE}
# Most inspections occur April through July (by district category)
inlet_recent_inspections |> 
  group_by(edit_month, Condition, district_category) |> 
  summarize(n=n()) |> 
  ggplot(aes(district_category, n, fill = Condition)) +
  geom_col() +
  theme_bw(10) + 
  facet_wrap(~edit_month, nrow=3) + 
  theme(axis.text.x=element_text(angle = 90)) +
  NULL

```
Analysis: Most inspections occur April through July.

Recent inspections by district category 
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(district_category) |> 
  count()|> 
  ggplot(aes(district_category, n)) +
  geom_col() +
  theme_bw(10) +
  theme(axis.text.x=element_text(angle = 90)) + 
  NULL

```
Analysis: The least number of inspections occur in residential districts.

Condition of recent inspections by district category
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(district_category, Condition) |> 
  count()|> 
  ggplot(aes(district_category, n, fill = Condition)) +
  geom_col() +
  theme_bw(10) +
  theme(axis.text.x=element_text(angle = 90))

```

Frame inspections by district category
```{r echo=FALSE}
inlet_recent_inspections |> 
  group_by(FRAME, district_category) |> 
  count() |> 
  ggplot(aes(n, district_category)) +
  geom_col() + 
  theme_bw(10) + 
  facet_wrap(~FRAME) + 
  NULL

```
Analysis: Most frame repair inspections occur in high-middle density districts.

Remaining questions/potential graphs: 
- Inlet condition by neighborhood
- Inlet condition by neighborhood, districts, and district categories
- When do inspections occur (month, year) by neighborhood, districts, and district categories?
- When do inspections occur (month, year) by neighborhood, districts, cleaning type, and sun_vs_shade? 

## F Parks
```{r echo=FALSE}
(parks <- st_read('input_data/NewHavenParksLayers_08032021/NewHavenParks.shp'))

parks |> mapview()

mapview(parks) +
  mapview(neighs, zcol = "Neighborhood", layer.name = "neighborhood") +
  mapview(inlet)

```

## G Parcels
```{r echo=FALSE}
```



