---
title: "New Haven GSI Stewardship"
author: "Clare Auld-Brokish (with Dexter Locke)"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---


# 0 set up
```{r include=FALSE}

# Load libraries 
packs <-c('tidyverse'   # cuz
          # , 'tidylog'   # prints out what was done in dplyr and tidyr; VERBOSE
          # , 'tidygeocoder' # geocode addressses
          , 'tidycensus'   # access Census data
          , 'sf'           # for spatial data support
          , 'mapview'      # web maps for zooming and panning around
          , 'janitor'   # helps clean things pipe-friendly cross-tabulations
          , 'leafsync'  # linked maps
          , 'magrittr'       # all of the pipes
          )

if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load the packages all at once
vapply(packs, library, character.only = TRUE, logical(1),
       logical.return = TRUE, quietly = TRUE)


```


# 1 Read in data
## A Neighborhoods
```{r include=FALSE}

neighs <- 
   st_read('input_data/New_Haven_Neighborhoods/Neighborhood.shp', as_tibble = TRUE) |> 
   select(Neighborhood = Neighbor_1) |> 
   st_transform(3857) %>%
   mutate(area_m2=as.double(st_area(.)))

```

```{r echo=FALSE}
neighs |> mapview(zcol = "area_m2")

```

## B Watersheds and Zoning
```{r include=FALSE}

sheds <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/12 digit Hydrologic Unit.geojson') |> 
  st_transform(st_crs(neighs))

```

```{r echo=FALSE}
mapview(neighs, alpha.regions = 0, lwd = 2) + mapview(sheds, zcol = 'Name') 

```

```{r include=FALSE}
zoning <- 
  st_read('../GSI STEWARDSHIP Project_data_too_big/_ags_data0AA96A07DF994B819C3A75306E2DD073.gdb') |> 
  st_transform(st_crs(neighs)) |> 
  select(-shape_Length,-shape_Area) %>%
  mutate(area_m2=as.double(st_area(.)))

```

```{r echo=FALSE}
zoning |> mapview(zcol = 
                    'districts'
                  )

```

```{r include=FALSE}
# Area in each zone
zoning |> 
  st_drop_geometry() |> 
  group_by(districts) |> 
  summarize(total_area=sum(area_m2)
            ,n=n()
            , mean_area=mean(area_m2)
            , median_area=median(area_m2)
            , sd_area=sd(area_m2)) |> 
  arrange(desc(total_area)) 

```

## C Parks
```{r include=FALSE}
parks <- st_read('input_data/NewHavenParksLayers_08032021/NewHavenParks.shp')

```

```{r echo=FALSE}
mapview(parks) +
  mapview(neighs, zcol = "Neighborhood", layer.name = "neighborhood")

```

# 2 Data Transformation
## A Inlet
Join neighborhoods, watersheds, and zoning in one dataset called Inlet (n = 290)
```{r include=FALSE}
inlet <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'inlet'
          , as_tibble = TRUE) |> 
  st_intersection(neighs) |> # adds neighborhood name
  st_intersection(sheds |> 
                    select(watershed = Name, HUC12)
                    ) |> 
  st_intersection(zoning |> 
                    select(zone_code, districts, district_category)
                  ) |> 
  mutate(  edit_year = lubridate::year(EditDate)
         , edit_month = lubridate::month(EditDate)
         , edit_date = lubridate::date(EditDate)
         , sun_vs_shade = str_extract(Planting_Plan, "^[^ ]+") |> str_to_title()
  ) 

```

```{r echo=FALSE}
mapviewOptions(
    legend.pos = 'bottomleft'
  , layers.control.pos = 'topright'
  )

```

View all inlets at once by neighborhood, zone, and watershed
``` {r echo=FALSE}
m_neigh <- inlet |> mapview(zcol = 'Neighborhood')
m_water <- inlet |> mapview(zcol = 'watershed')
m_zone  <- inlet |> mapview(zcol = 'districts')

leafsync::sync(m_neigh, m_water, m_zone)

```

### A1 Inlets per District 
``` {r include=FALSE}
n_gsi_district <- 
  inlet |> 
  group_by(districts) |> 
  summarize(  #total_area  = sum(area_m2)
            , n           = n()
            ) |> 
  st_drop_geometry()

zoning |> 
  group_by(districts) |> 
  summarise(district_area_m2 = sum(area_m2)) |> 
  left_join(n_gsi_district, by = 'districts') |> 
  mutate(gsi_per_district_area = ifelse(is.na(n), 0, n/district_area_m2))

```

### A2 Inlets per District Category 
``` {r echo=FALSE}
n_gsi_district_category <- 
  inlet |> 
  group_by(district_category) |> 
  summarize(  #total_area  = sum(area_m2)
            , n           = n()
            ) |> 
  st_drop_geometry()

```

### A3 Inlets per Watershed 
``` {r echo=FALSE}
n_gsi_watershed <- 
  inlet |> 
  group_by(watershed) |> 
  summarize(  #total_area  = sum(area_m2)
            , n           = n()
            ) |> 
  st_drop_geometry()

```

## B Inspection
Create a dataset with all inspection results from 2021 to 2025, beginning with condition (n = 1,256)
```{r include=FALSE}
inspection <- 
  st_read(  '../GSI STEWARDSHIP Project_data_too_big/50961b03-cb0b-49b9-a656-77d71bc30bbd.gdb'
          , 'GI_Inspection'
          , as_tibble = TRUE) |> 
mutate(Condition=factor(CONDITION,
    levels=c("Unknown", NA, "Very Poor", "Poor", "Fair", "Good", "Very Good", "Excellent"))
  , islegit=ifelse(Condition=="Unknown" | is.na(Condition)
                   , "notlegit"
                   , "legit"), 
  edit_year = lubridate::year(EditDate), 
    edit_month = lubridate::month(EditDate),
    edit_date = lubridate::date(EditDate)
  ) |>
  select(-EditDate) 

```

```{r echo=FALSE}
inspection |> 
  tabyl(CONDITION, islegit) |> 
  pivot_longer(-CONDITION, names_to = "islegit") |> 
  ggplot(aes(value, CONDITION, fill = CONDITION)) +
  geom_col() + 
  facet_wrap(~islegit) + 
  theme_linedraw(12) +
  labs(
    x = "Number of Inspections", 
    y = "Number of Inlets",
    title = "Most total inspections are legit"
  )

```

### B1 GSI per Neighborhood
Account for mean area of a neighborhood when comparing inlets across neighborhoods
```{r include=FALSE}
gsi_per_neigh <- 
  inlet |> 
  ungroup() |> 
  st_drop_geometry() |> 
  group_by(Neighborhood) |> 
  count()

# HelpMe_Dexter: is there a way to make the labels on these mapviews more helpful (more sensitive to smaller numbers or change the numbers? I added a inlets per neighborhood mapview to the Analysis Section 3.1 so maybe its redundant now here)

# multiply those really small numbers by 1000
# use the 'layer.name' argument more advantageously
# (UPDATE, that's what you did below)

neighs |> 
  left_join(gsi_per_neigh, by = 'Neighborhood') |> 
  mutate(gsi_per_neigh_area = ifelse(is.na(n), 0, 1000*(n / area_m2))) |> 
  mapview(zcol = 'gsi_per_neigh_area', layer.name = 'YOUR<br>NICE<br>NAME')

# HelpMe_Dexter: R says there is an error because "argument is of length zero". . . 
# I'm not getting that error, which is good, but I added the magrittr package
# which contains the '%>%' function
neighs |> 
  left_join(gsi_per_neigh, by = "Neighborhood") %>%
  mutate(gsi_per_neigh_area = ifelse(is.na(n), 0, 1000*(n / area_m2))) |> 
  # mutate(n = ifelse(is.na(n), 0, n)
  #        , area_m2 = st_area(.)
  #        , gsi_per_area = n / area_m2
  #        ) |> 
  mapview(zcol = 'gsi_per_neigh_area', layer.name = 'GSI counts /<br>neigh area<br>*1000')
  
```

### B2 Frequency of Inspections
```{r include=FALSE}
inspection |> map(~sum(is.na(.))) |> bind_rows() |> t()

inspection_freq <- 
  inspection |> 
  tabyl(ASSETGUID) |> 
  as_tibble() |> 
  arrange(desc(n))

```

## C Inspection_Long
Join inspections and inlets by FACILITYID and GlobalID to create a dataset that includes condition, neighborhood, watershed, and zone (n = 1,258)
```{r include=FALSE}
inspection_long <- 
  inspection |> 
  left_join(
    inlet |> select(-c(CreationDate, edit_date, Creator, Editor, edit_year, edit_month, GlobalID))
              , by = c('FACILITYID' = 'FACILITYID'))

```

## D Recent_Inspections
Includes only the most recently performed inspection for each inlet (n = 290)
```{r include=FALSE}
recent_inspections <- 
  inspection |> 
  group_by(ASSETGUID) |> 
  arrange(edit_date) |> 
  slice(1) |>  
  ungroup()

recent_inspections |> 
  tabyl(CONDITION) |> 
  as_tibble()
```

## E Inlet_Recent_Inspections
Combines neighborhood, watershed, and district data with recent inspections data (n = 290)
```{r include=FALSE}
inlet_recent_inspections <- 
  inlet |> 
  left_join(
    recent_inspections |> select(GlobalID, ASSETGUID, FACILITYID, DEBRIS, VEGETATION, OIL, FRAME, Condition, islegit)
    , by = c('GlobalID' = 'ASSETGUID'))

inlet_recent_inspections |> 
  st_drop_geometry() |> 
  group_by(Condition, islegit) |> 
  count()

```

# 3 Analysis 
## 1 Inlets per Watershed, Neighborhood, District, and District Category
```{r echo=FALSE}
n_gsi_watershed |> 
  st_drop_geometry() |> 
  group_by(n, watershed) |> 
  ggplot(aes(watershed, n, fill = n)) + 
  geom_col() + 
  theme_linedraw(12) + 
  labs (
    x = "Watershed",
    y = "Inlets",
    title = "Most inlets are located in the Cove River-Frontal LI Sound Watershed"
  )  

mapview(sheds, zcol = "Name", layer.name = "Watershed") +
  mapview(inlet)
  
gsi_per_neigh |> 
  st_drop_geometry() |> 
  group_by(n, Neighborhood) |> 
  ggplot(aes(Neighborhood, n, fill = n)) + 
  geom_col() + 
  theme_linedraw(12) + 
  theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = .5)) +
  labs (
    x = "Neighborhood",
    y = "Inlets",
    title = "Most inlets are located in the Hill Neighborhood"
  )

mapview(neighs, zcol = "Neighborhood", layer.name = "Neighborhood") +
  mapview(inlet)

n_gsi_district |> 
  st_drop_geometry() |> 
  group_by(n, districts) |> 
  ggplot(aes(n, districts, fill = n)) + 
  geom_col() + 
  theme_linedraw(12) + 
  labs (
    x = "Inlets",
    y = "Districts",
    title = "Residential and business districts have the highest number of inlets per district"
  )

mapview(zoning, zcol = "districts", layer.name = "Districts") +
  mapview(inlet)

n_gsi_district_category |> 
  st_drop_geometry() |> 
  group_by(n, district_category) |> 
  ggplot(aes(n, district_category, fill = n)) + 
  geom_col() + 
  theme_linedraw(12) + 
  labs (
    x = "Inlets",
    y = "District Categories",
    title = "High-middle density and general business districts have the highest number of inlets per district"
  )

mapview(zoning, zcol = "district_category", layer.name = "District Category") +
  mapview(inlet)

```

## 2 Focus of Inspections
```{r echo=FALSE}
inspection_long |> 
  group_by(watershed) |> 
  count() |> 
  ggplot(aes(watershed, n, fill =  n)) + 
  geom_col() + 
  theme_linedraw(12) +
  labs(
    x = "Watershed", 
    y = "Inspections",
    title = "Inspections focus primarily on the Cove River-Frontal LI Sound watershed"
  )

inspection_long |> 
  group_by(watershed, Neighborhood) |> 
  count() |> 
  ggplot(aes(n, Neighborhood, fill = n)) +
  geom_col() + 
  theme_linedraw(12) + 
  facet_wrap(~watershed) + 
  labs(
    x = "Inspections", 
    y = "Neighborhood",
    title = "Inspections focus primarily on the Cove River-Frontal LI Sound watershed in the Hill Neighborhood"
  )

inspection_long |> 
  group_by(districts) |> 
  count(districts) |> 
  ggplot(aes(districts, n, fill = n)) +
  geom_col() + 
  theme_linedraw(12) + 
  theme(axis.text.x=element_text(angle = 90)) +
  labs(
    x = "Districts", 
    y = "Inspections",
    title = "Most inspections occur in residential districts"
  )

```

## 3 Frequency of Inspections
```{r echo=FALSE}
inspection_freq |> 
  ggplot(aes(n, percent, fill = percent)) + 
  geom_col() +
  theme_linedraw(12) + 
  scale_x_binned() + 
  labs(
    y = "Percent",
    title = "The majority of inlets have been inspected roughly four times between 2021 and 2025"
  )

inspection_freq |> 
  ggplot(aes(n, valid_percent, fill = valid_percent)) + 
  geom_col() +
  theme_linedraw(12) + 
  scale_x_binned() + 
  labs(
    y = "Valid Percent",
    title = "The results for valid percent of inspections are similar to the results for percent (above)",
  )

```

Distribution of frequencies
```{r eval=FALSE, include=FALSE}
#HelpMe_Dexter: remind me what inspection_freq_freq was for? Is it still necessary to include given the above graphs? 

# how many times do you have 2 per, how many 3 per.. no need to keep

inspection_freq_freq <- 
  inspection_freq |> 
  tabyl(n) |> 
  as_tibble()

inspection_freq_freq |> 
  ggplot(aes(n,n_n)) + 
  geom_col() +
  theme_bw(16) +
  NULL 

```

## 4 Inspection Timing
```{r echo=FALSE}

recent_inspections |> 
  tabyl(Condition, islegit) |> 
  pivot_longer(-Condition, names_to = "islegit") |> 
  ggplot(aes(value, Condition, fill = Condition)) +
  geom_col() + 
  facet_wrap(~islegit) + 
  theme_linedraw(12) +
  labs(
    x = "Number of Inspections", 
    y = "Number of Inlets",
    title = "Most recent inspections are legit"
  )

recent_inspections |> 
  tabyl(edit_date)  |> 
  as_tibble() |> 
  ggplot(aes(edit_date, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth() + 
  theme_linedraw(12) +
  labs(
    x = "Last Inspection Year", 
    y = "Number of Inlets",
    title = "Most of the inlets were last inspected in 2021"
  )

recent_inspections |> 
  tabyl(edit_month)  |> 
  as_tibble() |> 
  ggplot(aes(edit_month, n)) +
  geom_point() + 
  geom_line() +
  geom_smooth() + 
  theme_linedraw(12) +
  scale_x_binned() + 
  labs(
    x = "Last Inspection Month", 
    y = "Number of Inlets",
    title = "Most of the inlets were last inspected between April and June"
  )

inspection_long |> 
  group_by(edit_month, Condition) |> 
  summarize(n=n()) |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_linedraw(12) +  
  facet_wrap(~edit_month, nrow=3) + 
  labs(
    x = "Inspections", 
    y = "Condition",
    title = "There is minimal positive correlation between inspection month and inlet condition",
    subtitle = "Inlets inspected in January and December are consistently in fair condition"
  )
```

## 5 Planting Plan (sun_vs_shade)
```{r echo=FALSE}
inlet |>
  st_drop_geometry() |> 
  tabyl(Planting_Plan, sun_vs_shade)

inlet |> 
  st_drop_geometry() |>
  tabyl(sun_vs_shade) |> 
  ggplot(aes(sun_vs_shade, percent, fill = sun_vs_shade)) +
  geom_col() + 
  theme_linedraw(12) + 
  labs(
    x = "Planting Plan", 
    y = "Percent", 
    title = "There are more sun planting plans than shade planting plans for inlets across the city",
    subtitle = "About 46 percent of inlets don't have a planting plan"
  )

inlet |> 
  st_drop_geometry() |> 
  group_by(sun_vs_shade, Assigned, watershed) |> 
  ggplot(aes(sun_vs_shade, fill = Assigned)) + 
  geom_bar() + # changed to geom_bar from geom_col
  theme_linedraw() + 
  theme(axis.text.y = element_blank(), legend.position = 'bottom') + 
  facet_grid(~str_wrap(watershed, 25)) + # wrapping
  labs(
    x = "Planting Plan",
    title = "The Cove River-Frontal Long Island Sound Watershed has the largest number of planting plans",
    subtitle = "Most are NA and assigned routine cleaning"
  )

inlet |> 
  st_drop_geometry() |> 
  group_by(sun_vs_shade, Assigned, Neighborhood) |> 
  ggplot(aes(sun_vs_shade, Assigned, fill = Assigned)) + 
  geom_col()+ 
  theme_linedraw(12) + 
  theme(axis.text.y = element_blank()) + 
  facet_wrap(~Neighborhood) +
  labs(
    x = "Planting Plan",
    title = "The Hill neighborhood has the largest number of planting plans",
    subtitle = "Most planting plans are NA and assigned routine cleaning"
  )
  
inlet |> 
  st_drop_geometry() |> 
  group_by(sun_vs_shade, Assigned, districts) |> 
  ggplot(aes(sun_vs_shade, Assigned, fill = Assigned)) + 
  geom_col()+ 
  theme_linedraw(12) + 
  theme(axis.text.y = element_blank()) + 
  facet_wrap(~districts) +
  labs(
    x = "Planting Plan",
    title = "Business and residential districts have the largest number of planting plans",
    subtitle = "Many inlets in these districts are not assigned either emergency or routine cleaning"
  )
  
```

## 6 Inlet Condition
### 1 Inlet Condition
```{r echo=FALSE}
recent_inspections |> 
  tabyl(Condition) |> 
  ggplot(aes(Condition, n, fill = Condition)) +
  geom_col() + 
  theme_linedraw(12) +
  labs(
    y = "Inlets (number)", 
    title = "Most inlets are in Poor/Fair condition"
  )

recent_inspections |> 
  tabyl(Condition) |> 
  ggplot(aes(Condition, percent, fill = Condition)) +
  geom_bar(stat="identity", width =1) +
  coord_polar("y", start=0) +
  theme_linedraw(12) +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank()
    ) + 
  labs(
    y = "Inlets (percent)", 
    title = "Most inlets are in Poor/Fair condition"
  )

```

### 2 Assigned Cleanings
```{r echo=FALSE}
inspection_long |> 
  tabyl(Assigned) |> 
  as_tibble()

inspection_long |> 
  tabyl(Neighborhood, Assigned) |> 
  as_tibble()

inspection_long |> 
  tabyl(Assigned) |> 
  ggplot(aes(n, Assigned, fill = n)) +
  geom_col() + 
  theme_linedraw(12) + 
  labs(
    x = "Inlets", 
    y = "Cleaning Assignment",
    title = "Most inlets are assigned emergency cleaning"
  )

inspection_long |> 
  group_by(edit_year, Assigned) |> 
  summarize(n=n()) |> 
  ggplot(aes(Assigned, n, fill = Assigned)) +
  geom_col() + 
  theme_linedraw(12) +
  theme(axis.text.x=element_text(angle = 90)) +
  facet_wrap(~edit_year) + 
  labs(
    x = "Cleaning Assignment",
    y = "Inlets",
    title = "The number of emergency cleanings was greatest from 2021-2023", 
    subtitle = "This may correlate with inconsistent maintenance during these years"
  )

inspection_long |> 
  group_by(Assigned, Condition) |> 
  count('Routine Cleaning') |> 
  ggplot(aes(n, Condition, fill = Condition)) + 
  geom_col() + 
  facet_wrap(~Assigned, scales="free") +
  theme_linedraw(12) + 
  labs(
    x = "Inspections", 
    y = "Condition",
    title = "Cleaning assignment does not measurably impact inlet condition"
  )

inspection_long |> 
  group_by(Neighborhood, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~Neighborhood) + 
  theme_linedraw(12) + 
  labs(
    x = "Inspections", 
    y = "Condition",
    title = "Inlets in the Hill Neighborhood have the highest number of inspections and emergency cleanings"
  )

inspection_long |> 
  group_by(watershed, Assigned, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Assigned)) +
  geom_col() + 
  theme_linedraw() +
  facet_wrap(~watershed) + 
  theme_linedraw(12) + 
  labs(
    x = "Inspections", 
    y = "Condition",
    title = "Inlets in the Cove River-Frontal LI Sound watershed have the highest number of inspections and emergency cleanings"
  )

```

### 3 Presence of Vegetation, Debris, Oil; Frame Condition
```{r include=FALSE}
recent_inspections |> 
  tabyl(DEBRIS) |> 
  as_tibble() |> 
  arrange(desc(n))

```

```{r echo=FALSE}
recent_inspections |> 
  group_by(VEGETATION, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = VEGETATION)) +
  geom_col() + 
  theme_linedraw(12) +
  theme(
    axis.ticks.y = element_blank()
  ) +
  labs(
    x = "Condition", 
    y = "Inlets",
    title = "Most inlets did not have vegetation during their most recent inspection"
  )

recent_inspections |> 
  group_by(Condition, OIL) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = OIL)) +
  geom_col() + 
  theme_linedraw(12) +
  theme(
    axis.ticks.y = element_blank()
  ) +
  labs(
    x = "Condition", 
    y = "Inlets",
    title = "Most inlets did not have oil during their most recent inspection"
  )

recent_inspections |> 
  group_by(DEBRIS, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = DEBRIS)) + 
  geom_col() + 
  theme_linedraw(12) +
  theme(
    axis.ticks.y = element_blank()
  ) +
  labs(
    x = "Number of Inlets", 
    title = "Most inlets had debris during their most recent inspection"
  )

recent_inspections |> 
  group_by(FRAME, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  facet_wrap(~FRAME) + 
  theme_linedraw(12) +
  theme(
    axis.ticks.y = element_blank()
  ) +
  labs(
    x = "Inlets", 
    title = "Only a few inlets needed a frame repair"
  )


inlet_recent_inspections |> 
  group_by(FRAME, district_category, Condition) |> 
  count() |> 
  ggplot(aes(n, district_category, fill = Condition)) +
  geom_col() + 
  theme_linedraw(12) + 
  facet_wrap(~FRAME) + 
  labs(
    x = "Inlets",
    y = "District Category", 
    title = "Frame inspections by district category and condition",
    subtitle = "The highest number of frame repairs occurs in high-middle density districts"
  )

```

### 4 By Watershed, Neighborhood, District, and District Category
```{r echo=FALSE}
inspection_long |> 
  ggplot(aes(Condition, fill = Neighborhood)) +
  geom_bar() +
  facet_wrap(~watershed, scales = "free_y") + 
  theme_linedraw(12) + 
  theme(axis.text.x=element_text(angle = 90)) +
  labs(
    x = "Condition", 
    y = "Inspections",
    title = "Inlet condition by watershed and neighborhood"
  )

inlet_recent_inspections |> 
  group_by(watershed, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_linedraw(12) +
  theme(legend.position = "bottom") +
  facet_wrap(~watershed) +
  labs(
    x = "Inlets", 
    title = "Most inlets in the Cove River-Frontal Long Island Sound and West River Watersheds were in fair/poor condition", 
    subtitle = "Not enough inlets in other watersheds to determine if watershed and inlet condition are causally linked"
  )

inlet_recent_inspections |> 
  group_by(Neighborhood, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() +
  theme_linedraw(12) +
  theme(legend.position = "bottom") +
  facet_wrap(~Neighborhood) +
  labs(
    x = "Inlets", 
    title = "Most inlets are located in the Dwight and Hill neighborhoods", 
    subtitle = "Not enough inlets in other neighborhoods to determine if neighborhood and inlet condition are causally linked"
  )

inlet_recent_inspections |> 
  group_by(districts, Condition) |> 
  count() |> 
  ggplot(aes(n, Condition, fill = Condition)) +
  geom_col() + 
  theme_linedraw(12) +
  theme(legend.position = "bottom") +
  facet_wrap(~districts) +
  labs(
    x = "Inlets", 
    title = "Most inlets in business and residential districts were in fair/poor condition", 
    subtitle = "Not enough inlets in other districts to determine if district and inlet condition are causally linked"
  )

inlet_recent_inspections |> 
  group_by(watershed, districts, Condition) |> 
  count() |> 
  ggplot(aes(Condition, n, fill = districts)) +
  geom_col() + 
  facet_wrap(~watershed) +
  theme_linedraw(12) + 
  theme(axis.text.x=element_text(angle = 90)) + 
  labs(
    y = "Inlets", 
    title = "Inlet condition by watershed and district",
    subtitle = "Most inlets were located in business districts in the Cove River-Frontal LI Sound Watershed"
  )

inlet_recent_inspections |> 
  group_by(district_category, Condition) |> 
  count()|> 
  ggplot(aes(district_category, n, fill = Condition)) +
  geom_col() +
  theme_linedraw(12) +
  theme(axis.text.x=element_text(angle = 90)) +
  labs(
    x = "District Category",
    y = "Inlets", 
    title = "Condition of recent inspections by district category",
    subtitle = "Inlets in very good condition are found in general business and planned development districts"
  )

```